Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Subjective Wellbeing

### 2019-11-01 Dissonance_Santiago_Title_Page_v1.Rmd
Added discussion based on the files sent by Beatriz, and prepared files for submission

---
title: Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Subjective Wellbeing
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
  Transportation in most of the world has been dominated for decades by a fascination with the automobile. Nonetheless, there is an increasing recognition that to achieve a variety of economic, environmental, and public health policy goals it is important to attract and retain active travelers and users of public transportation. A challenge with the way transportation policy is developed, however, is that it tends to focus on utilitarian considerations, which may miss other potential aspects of transportation that users value. In particular, subjective wellbeing (SWB) has been proposed as a way to enhance our understanding of the preferences and choices of travellers, as well as a way to evaluate the benefits of transportation beyond utilitarian considerations. The objective of this paper is to analyze the modes that people commonly use, and to what extent they are aligned (or not) with a variety of affective values associated with SWB. In other words, we are interested in the potential for dissonance with respect to the primary mode of travel, from the perspective of affective values. The study is based on data collected from a sample of travellers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name the mode or modes that they associate with the affective values of freedom, enjoyment, happiness, poverty, luxury and status. The results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, public transportation is most commonly associated with poverty, and the automobil is most commonly associated with luxury and status.

@@

<!-- This paper is an R markdown document that includes self-contained, reproducible analysis. To reproduce the paper the following files are needed:

1. Dissonance_Santiago_v1.Rmd (this R markdown document)

2. `Dissonance_Santiago.RData`. For reference: this data file was created using R notebook `00. Analysis of Dissonance Detailed.Rmd`. The initial analysis was done in the same R notebook. Neither of these two notebooks are needed for this document.

3. For compiling into a pdf in the style of an Elsevier document: elsarticle.cls, elsevier-harvard.csl, elsevier-harvard-without-titles.csl, elsevier-vancouver.csl, elsevier-without-titles.cls, elsevier-with-titles.csl, elsevier-with-titles-alphabetical.csl, mybibfile.bib
-->

```{r load-packages, include=FALSE}

# Load packages used in this document

library(rticles)
library(readxl)
library(tidyverse)
library(rgdal)
library(sf)
library(ggmosaic)
library(ggthemes)
library(kableExtra)
library(gridExtra)
```

```{r load-data, include=FALSE}

# To run this notebook, the file `Dissonance_Santiago.RData` is needed. This file contains the following object:
# 
# master: a dataframe with information collected from 451 respondents. The variables in the dataframe are:
# 
# - NUMERO: Respondent ID
# - USED: Main mode of transportation used by the respondent
# - FREEDOM: Mode(s) that the respondent associates with feelings of freedom
# - ENJOYMENT: Mode(s) that the respondent associates with feelings of enjoyment
# - HAPPINESS: Mode(s) that the respondent associates with feelings of happiness
# - POVERTY: Mode(s) that the respondent associates with feelings of poverty
# - LUXURY: Mode(s) that the respondent associates with feelings of luxury
# - STATUS: Mode(s) that the respondent associates with feelings of status
# - SAFETY: Mode(s) that the respondent associates with feelings of safety
# - TIME: Mode(s) that the respondent associates with feelings of WASTED TIME
# - EFFICIENCY: Mode(s) that the respondent associates with feelings of efficiency
# - AGE: Age of the respondent
# - EDUCATION: Level of education of the respondent
# - INCOME: Level of income of respondent
# - TRAVEL_TIME: Trip duration of regular commuting trip 
# - WEIGHT_*: A weight for the responses by ATTRIBUTE. Some respondents selected more than one mode that they associated with feelings of {FREEDOM, ENJOYMENT, HAPPINESS, POVERTY, LUXURY, STATUS, SAFETY, TIME, and EFFICIENCY}. To avoid giving greater weight to respondents who indicated several modes of transportation, a weight was calculated that is the inverse of the number of modes indicated. These weights can be used to weight the frequencies prior to any analysis.
# 
# The modes of transportation (USED, FREEDOM, ENJOYMENT, HAPPINESS, POVERTY, LUXURY, STATUS, SAFETY, TIME, EFFICIENCY) are coded as: 
# "A": Car
# "B": Taxi
# "C": Colectivo
# "D": Moto
# "E": Metro
# "F": Bus
# "G": Bicycle
# "H": Walk
#
# AGE is coded as:
#
# "A": Less18
# "B": 18to24
# "C": 25to34
# "D": 35to54
# "E": 55to64
# "F": 65more
# 
# EDUCATION is coded as:
# 
# "A": ELEMENTARY
# "B": SECUNDARY
# "C": PROFESS_TECH
# "D": COLLEGE
# "E": POSTGRAD
#
# INCOME is coded as (in pesos?):
# 
# "A": Less423
# "B": 423to639
# "C": 639to977
# "D": 977to1550
# "E": 1550to2380
# "F": 2380more
# 
# TRAVEL_TIME is coded as:
#
# "A": 0-20 min
# "B": 20-40 min
# "C": 40-60 min
# "D": 1h and more

# Read the RData file:

load("Dissonance_Santiago.RData")
```

```{r data-preparation, include=FALSE}

# The master dataframe is processed for analysis. We plan to conduct two distinct types of analysis:
#
# 1. Who experiences dissonace?
#
# Analysis of dissonace using the characteristics of respondents (i.e., AGE, EDUCATION, INCOME, (Mode) USED, and TRAVEL_TIME) by ATTRIBUTE (i.e., FREEDOM et al.). In this case, dissonance is defined as follows:
#
# >> If the mode used is identical to the mode that the respondent associates with ATTRIBUTE (i.e. FREEDOM et al.), we consider that there is no dissonance; otherwise, we consider that there is dissonance. For example, if the respondent's main mode of transportation is Car, and the respondent associates Car with feelings of FREEDOM, then there is no dissonance. On the other hand, if the respondent's main mode of transportation is Car, but associates Walking with feelings of FREEDOM, there is dissonance.
# 
# Given this, we will calculate how frequently there is dissonance at different levels of AGE, EDUCATION, INCOME, MODE, TRAVEL_TIME. Tests of independence can be implemented.
#
# 2. For those respondents who experience dissonance, what mode(s) are better aligned with various affective values
#
# Analysis of dissonance using the mode USED and the mode respondents associated with ATTRIBUTE, then stratified by characteristics of respondents (i.e., AGE et al.) in this case, we see how frequently each mode was associated with ATTRIBUTE, given the main mode of transportation of the respondents.
#
# Here the dataframes are prepared to support these two types of analysis

#####################################################################################
# Prepare a dataframe for Analysis 1: Who experiences dissonance by affective value #
#####################################################################################

# First, select the variables needed from the master table, and code dissonance as 1 (no dissonance) or two (dissonance):

Overall_dissonance <- master %>%
  transmute(USED, 
            FREEDOM_DISSONANCE = ifelse(USED == FREEDOM, 1, 2),
            ENJOYMENT_DISSONANCE = ifelse(USED == ENJOYMENT, 1, 2),
            HAPPINESS_DISSONANCE = ifelse(USED == HAPPINESS, 1, 2),
            POVERTY_DISSONANCE = ifelse(USED == POVERTY, 1, 2),
            LUXURY_DISSONANCE = ifelse(USED == LUXURY, 1, 2),
            STATUS_DISSONANCE = ifelse(USED == STATUS, 1, 2),
            SAFETY_DISSONANCE = ifelse(USED == SAFETY, 1, 2),
            TIME_DISSONANCE = ifelse(USED == TIME, 1, 2),
            EFFICIENCY_DISSONANCE = ifelse(USED == EFFICIENCY, 1, 2),
            AGE,
            EDUCATION,
            INCOME,
            TRAVEL_TIME)

# Recode variables and convert to factors:

Overall_dissonance <- Overall_dissonance %>%
  mutate(USED = case_when(USED == "A" ~ "Car",
                                  USED == "B" | USED == "C" | USED == "D" ~ "Other",
                                  USED == "E" | USED == "F" ~ "Public",
                                  USED == "G" | USED == "H" ~ "Active"),
         AGE = case_when(AGE == "A" | AGE == "B" | AGE == "C" ~ "< 35",
                         AGE == "D" ~ "35 - 54",
                         AGE == "E" | AGE == "F" ~ "> 54"),
         EDUCATION = case_when(EDUCATION == "A" | EDUCATION == "B" ~ "K-12",
                               EDUCATION == "C" | EDUCATION == "D" ~ "Tech/Univ",
                               EDUCATION == "E" ~ "Graduate"),
         INCOME = case_when(INCOME == "A" | INCOME == "B" ~ "Low",
                            INCOME == "C" | INCOME == "D" ~ "Middle",
                            INCOME == "E" | INCOME == "F" ~ "High"),
         TRAVEL_TIME = case_when(TRAVEL_TIME == "A" ~ "< 20 min",
                                 TRAVEL_TIME == "B" ~ "20 - 40 min",
                                 TRAVEL_TIME == "C" ~ "40 - 60 min",
                                 TRAVEL_TIME == "D" ~ "> 60 min")) %>%
  mutate(USED = factor(USED, levels = c("Car", "Active", "Public", "Other"), labels = c("Car", "Active", "Public", "Other")),
         FREEDOM_DISSONANCE = factor(FREEDOM_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         ENJOYMENT_DISSONANCE = factor(ENJOYMENT_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         HAPPINESS_DISSONANCE = factor(HAPPINESS_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         POVERTY_DISSONANCE = factor(POVERTY_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         LUXURY_DISSONANCE = factor(LUXURY_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         STATUS_DISSONANCE = factor(STATUS_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         SAFETY_DISSONANCE = factor(SAFETY_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         TIME_DISSONANCE = factor(TIME_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         EFFICIENCY_DISSONANCE = factor(EFFICIENCY_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         AGE = factor(AGE, levels = c("< 35", "35 - 54", "> 54"), labels = c("< 35", "35 - 54", "> 54")),
         EDUCATION = factor(EDUCATION, levels = c("K-12", "Tech/Univ", "Graduate"), labels = c("K-12", "Tech/Univ", "Graduate")),
         INCOME = factor(INCOME, levels = c("Low", "Middle", "High"), labels = c("Low", "Middle", "High")),
         TRAVEL_TIME = factor(TRAVEL_TIME, levels = c("< 20 min", "20 - 40 min", "40 - 60 min", "> 60 min"), labels = c("< 20 min", "20 - 40 min", "40 - 60 min", "> 60 min")))

###############################################################################################
# Prepare a dataframe for Analysis 2: What mode(s) align better with various affective values #
###############################################################################################

# First, split the responses regarding the modes associated with each attribute, so that we can record each mode selected individually:

Modes_by_Attribute <- master %>% 
    mutate(FREEDOM = strsplit(as.character(FREEDOM), ""),
           ENJOYMENT = strsplit(as.character(ENJOYMENT), ""),
           HAPPINESS = strsplit(as.character(HAPPINESS), ""),
           POVERTY = strsplit(as.character(POVERTY), ""),
           LUXURY = strsplit(as.character(LUXURY), ""),
           STATUS = strsplit(as.character(STATUS), ""),
           SAFETY = strsplit(as.character(SAFETY), ""),
           TIME = strsplit(as.character(TIME), ""),
           EFFICIENCY = strsplit(as.character(EFFICIENCY), ""))

# Here, we convert each of the responses into an individual row and stack the resulting dataframes, after adding the name of the attribute for identification:

Modes_by_Attribute <- rbind(data.frame(Modes_by_Attribute %>% 
                                         select(-c(ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY, 
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY, 
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>% 
                                         rename(IDEAL = FREEDOM, WEIGHT = WEIGHT_FREEDOM) %>%
                                         unnest(IDEAL),
                                       ATTRIBUTE = "FREEDOM"),
                            data.frame(Modes_by_Attribute %>%
                                         select(-c(FREEDOM, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY, 
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY, 
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>% 
                                         rename(IDEAL = ENJOYMENT, WEIGHT = WEIGHT_ENJOYMENT) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "ENJOYMENT"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY, 
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM,
                                                   WEIGHT_ENJOYMENT,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY, 
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>% 
                                         rename(IDEAL = HAPPINESS, WEIGHT = WEIGHT_HAPPINESS) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "HAPPINESS"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY,
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY,
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>%
                                         rename(IDEAL = POVERTY, WEIGHT = WEIGHT_POVERTY) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "POVERTY"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   #LUXURY,
                                                   STATUS,
                                                   SAFETY,
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   #WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY,
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>%
                                         rename(IDEAL = LUXURY, WEIGHT = WEIGHT_LUXURY) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "LUXURY"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   #STATUS,
                                                   SAFETY,
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   #WEIGHT_STATUS,
                                                   WEIGHT_SAFETY,
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>%
                                         rename(IDEAL = STATUS, WEIGHT = WEIGHT_STATUS) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "STATUS"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>%
                                         rename(IDEAL = SAFETY, WEIGHT = WEIGHT_SAFETY) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "SAFETY"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY,
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY,
                                                   WEIGHT_EFFICIENCY)) %>%
                                         rename(IDEAL = TIME, WEIGHT = WEIGHT_TIME) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "TIME"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY,
                                                   TIME, 
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY,
                                                   WEIGHT_TIME)) %>%
                                         rename(IDEAL = EFFICIENCY, WEIGHT = WEIGHT_EFFICIENCY) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "EFFICIENCY"))

# There are some missing values. Drop all NAs:

Modes_by_Attribute <- drop_na(Modes_by_Attribute)

# We now recode the variables:

Modes_by_Attribute <- Modes_by_Attribute %>%
  mutate(USED = case_when(USED == "A" ~ "Car",
                                  USED == "B" | USED == "C" | USED == "D" ~ "Other",
                                  USED == "E" | USED == "F" ~ "Public",
                                  USED == "G" | USED == "H" ~ "Active"),
         IDEAL = case_when(IDEAL == "A" ~ "Car",
                                  IDEAL == "B" | IDEAL == "C" | IDEAL == "D" ~ "Other",
                                  IDEAL == "E" | IDEAL == "F" ~ "Public",
                                  IDEAL == "G" | IDEAL == "H" ~ "Active"),
         AGE = case_when(AGE == "A" | AGE == "B" | AGE == "C" ~ "< 35",
                         AGE == "D" ~ "35 - 54",
                         AGE == "E" | AGE == "F" ~ "> 54"),
         EDUCATION = case_when(EDUCATION == "A" | EDUCATION == "B" ~ "K-12",
                               EDUCATION == "C" | EDUCATION == "D" ~ "Tech/Univ",
                               EDUCATION == "E" ~ "Graduate"),
         INCOME = case_when(INCOME == "A" | INCOME == "B" ~ "Low",
                            INCOME == "C" | INCOME == "D" ~ "Middle",
                            INCOME == "E" | INCOME == "F" ~ "High"),
         TRAVEL_TIME = case_when(TRAVEL_TIME == "A" ~ "< 20 min",
                                 TRAVEL_TIME == "B" ~ "20 - 40 min",
                                 TRAVEL_TIME == "C" ~ "40 - 60 min",
                                 TRAVEL_TIME == "D" ~ "> 60 min"))

# After recoding the , there will be some duplicate entries, because a person may have attributed, say, feelings of freedom to walking and cycling, so in the table this will appear as two instances of "Active". This is ok, because each is weighted appropriately.

# Convert to factors:

Modes_by_Attribute <- Modes_by_Attribute %>%
  mutate(USED = factor(USED, levels = c("Car", "Active", "Public", "Other"), labels = c("Car", "Active", "Public", "Other")),
         IDEAL = factor(IDEAL, levels = c("Car", "Active", "Public", "Other"), labels = c("Car", "Active", "Public", "Other")),
         AGE = factor(AGE, levels = c("< 35", "35 - 54", "> 54"), labels = c("< 35", "35 - 54", "> 54")),
         EDUCATION = factor(EDUCATION, levels = c("K-12", "Tech/Univ", "Graduate"), labels = c("K-12", "Tech/Univ", "Graduate")),
         INCOME = factor(INCOME, levels = c("Low", "Middle", "High"), labels = c("Low", "Middle", "High")),
         TRAVEL_TIME = factor(TRAVEL_TIME, levels = c("< 20 min", "20 - 40 min", "40 - 60 min", "> 60 min"), labels = c("< 20 min", "20 - 40 min", "40 - 60 min", "> 60 min")))
```

```{r cross-tabs-function, include=FALSE}

## Define functions for crosstabs; these functions enhance the base crosstabs functions and make it easier to report cross tabulations in different ways.

crosstab <- function (..., dec.places = NULL,
                           type = NULL,
                           style = "wide",
                           row.vars = NULL,
                           col.vars = NULL,
                           percentages = TRUE, 
                           addmargins = TRUE,
                           subtotals=TRUE)

###################################################################################
#                                                                                 #
# Function created by Dr Paul Williamson, Dept. of Geography and Planning,        #
# School of Environmental Sciences, University of Liverpool, UK.                  #
#                                                                                 #
# Adapted from the function ctab() in the catspec packge.                         #
#                                                                                 #
# Version: 12th July 2013                                                         #
#                                                                                 #
# Output best viewed using the companion function print.crosstab()                #
#                                                                                 #
###################################################################################

  #Declare function used to convert frequency counts into relevant type of proportion or percentage
  {mk.pcnt.tbl <- function(tbl, type) {
        a <- length(row.vars)
        b <- length(col.vars)
        mrgn <- switch(type, column.pct = c(row.vars[-a], col.vars), 
                            row.pct = c(row.vars, col.vars[-b]),
                            joint.pct = c(row.vars[-a], col.vars[-b]),
                            total.pct = NULL)
        tbl <- prop.table(tbl, mrgn)
        if (percentages) {
            tbl <- tbl * 100
        }
        tbl
    }

    #Find no. of vars (all; row; col) for use in subsequent code
    n.row.vars <- length(row.vars)
    n.col.vars <- length(col.vars)
    n.vars <- n.row.vars + n.col.vars

    #Check to make sure all user-supplied arguments have valid values
    stopifnot(as.integer(dec.places) == dec.places, dec.places > -1)
    #type: see next section of code
    stopifnot(is.character(style))    
    stopifnot(is.logical(percentages))
    stopifnot(is.logical(addmargins))
    stopifnot(is.logical(subtotals))
    stopifnot(n.vars>=1)

    #Convert supplied table type(s) into full text string (e.g. "f" becomes "frequency")
    #If invalid type supplied, failed match gives user automatic error message
    types <- NULL
    choices <- c("frequency", "row.pct", "column.pct", "joint.pct", "total.pct")
    for (tp in type) types <- c(types, match.arg(tp, choices))
    type <- types

    #If no type supplied, default to 'frequency + total' for univariate tables and to
    #'frequency' for multi-dimenstional tables

    #For univariate table....
    if (n.vars == 1) {
      if (is.null(type)) {
        # default = freq count + total.pct  
        type <- c("frequency", "total.pct")
        #row.vars <- 1
      } else {
        #and any requests for row / col / joint.pct must be changed into requests for 'total.pct'
        type <- ifelse(type == "frequency", "frequency", "total.pct")
      }
    #For multivariate tables...
    } else if (is.null(type)) {
      # default = frequency count  
      type <- "frequency"
    }

    #Check for integrity of requested analysis and adjust values of function arguments as required

    if ((addmargins==FALSE) & (subtotals==FALSE)) {
      warning("WARNING: Request to suppress subtotals (subtotals=FALSE) ignored because no margins requested (addmargins=FALSE)")
      subtotals <- TRUE
    }

    if ((n.vars>1) & (length(type)>1) & (addmargins==TRUE)) {
      warning("WARNING: Only row totals added when more than one table type requested")
      #Code lower down selecting type of margin implements this...
    }

    if ((length(type)>1) & (subtotals==FALSE)) { 
      warning("WARNING: Can only request supply one table type if requesting suppression of subtotals; suppression of subtotals not executed")
      subtotals <- TRUE
    }

    if ((length(type)==1) & (subtotals==FALSE)) {
      choices <- c("frequency", "row.pct", "column.pct", "joint.pct", "total.pct")
      tp <- match.arg(type, choices)
      if (tp %in% c("row.pct","column.pct","joint.pct")) {
          warning("WARNING: subtotals can only be suppressed for tables of type 'frequency' or 'total.pct'")
          subtotals<- TRUE
         }
    }

    if ((n.vars > 2) & (n.col.vars>1) & (subtotals==FALSE)) 
      warning("WARNING: suppression of subtotals assumes only 1 col var; table flattened accordingly")

    if ( (subtotals==FALSE) & (n.vars>2) )  {
    #If subtotals not required AND total table vars > 2
    #Reassign all but last col.var as row vars
    #[because, for simplicity, crosstabs assumes removal of subtotals uses tables with only ONE col var]
    #N.B. Subtotals only present in tables with > 2 cross-classified vars...
      if (length(col.vars)>1) {
        row.vars <- c(row.vars,col.vars[-length(col.vars)])
        col.vars <- col.vars[length(col.vars)]
        n.row.vars <- length(row.vars)
        n.col.vars <- 1
      }
    }

    #If dec.places not set by user, set to 2 unlesss only one table of type frequency requested,
    #in which case set to 0.  [Leaves user with possibility of having frequency tables with > 0 dp]
    if (is.null(dec.places)) {
      if ((length(type)==1) & (type[1]=="frequency")) {
        dec.places <- 0
      } else {
        dec.places <-2
      }
    }

    #Take the original input data, whatever form originally supplied in,
    #convert into table format using requested row and col vars, and save as 'tbl'

    args <- list(...)

    if (length(args) > 1) {
        if (!all(sapply(args, is.factor))) 
            stop("If more than one argument is passed then all must be factors")
        tbl <- table(...)
    }
    else {
        if (is.factor(...)) {
            tbl <- table(...)
        }
        else if (is.table(...)) {
            tbl <- eval(...)
        }
        else if (is.data.frame(...)) {
            #tbl <- table(...)
            if (is.null(row.vars) && is.null(col.vars)) {
              tbl <- table(...)
            }
            else {
              var.names <- c(row.vars,col.vars)
              A <- (...)
              tbl <- table(A[var.names])
              if(length(var.names==1)) names(dimnames(tbl)) <- var.names
              #[table() only autocompletes dimnames for multivariate crosstabs of dataframes]
            }
        }
        else if (class(...) == "ftable") {
            tbl <- eval(...)
            if (is.null(row.vars) && is.null(col.vars)) {
                row.vars <- names(attr(tbl, "row.vars"))
                col.vars <- names(attr(tbl, "col.vars"))
            }
            tbl <- as.table(tbl)
        }
        else if (class(...) == "ctab") {
            tbl <- eval(...)
            if (is.null(row.vars) && is.null(col.vars)) {
                row.vars <- tbl$row.vars
                col.vars <- tbl$col.vars
            }
            for (opt in c("dec.places", "type", "style", "percentages", 
                "addmargins", "subtotals")) if (is.null(get(opt))) 
                assign(opt, eval(parse(text = paste("tbl$", opt, 
                  sep = ""))))
            tbl <- tbl$table
        }
        else {
            stop("first argument must be either factors or a table object")
        }
    }

    #Convert supplied table style into full text string (e.g. "l" becomes "long")
    style <- match.arg(style, c("long", "wide"))

    #Extract row and col names to be used in creating 'tbl' from supplied input data
    nms <- names(dimnames(tbl))
    z <- length(nms)
    if (!is.null(row.vars) && !is.numeric(row.vars)) {
        row.vars <- order(match(nms, row.vars), na.last = NA)
    }
    if (!is.null(col.vars) && !is.numeric(col.vars)) {
        col.vars <- order(match(nms, col.vars), na.last = NA)
    }
    if (!is.null(row.vars) && is.null(col.vars)) {
        col.vars <- (1:z)[-row.vars]
    }
    if (!is.null(col.vars) && is.null(row.vars)) {
        row.vars <- (1:z)[-col.vars]
    }
    if (is.null(row.vars) && is.null(col.vars)) {
        col.vars <- z
        row.vars <- (1:z)[-col.vars]
    }

    #Take the original input data, converted into table format using supplied row and col vars (tbl)
    #and create a second version (crosstab) which stores results as percentages if a percentage table type is requested.
    if (type[1] == "frequency") 
      crosstab <- tbl
    else 
      crosstab <- mk.pcnt.tbl(tbl, type[1])

    #If multiple table types requested, create and add these to 
    if (length(type) > 1) {
      tbldat <- as.data.frame.table(crosstab)
      z <- length(names(tbldat)) + 1
      tbldat[z] <- 1
      pcntlab <- type
      pcntlab[match("frequency", type)] <- "Count"
      pcntlab[match("row.pct", type)] <- "Row %"
      pcntlab[match("column.pct", type)] <- "Column %"
      pcntlab[match("joint.pct", type)] <- "Joint %"
      pcntlab[match("total.pct", type)] <- "Total %"
      for (i in 2:length(type)) {
          if (type[i] == "frequency") 
              crosstab <- tbl
          else crosstab <- mk.pcnt.tbl(tbl, type[i])
          crosstab <- as.data.frame.table(crosstab)
          crosstab[z] <- i
          tbldat <- rbind(tbldat, crosstab)
      }
      tbldat[[z]] <- as.factor(tbldat[[z]])
      levels(tbldat[[z]]) <- pcntlab
      crosstab <- xtabs(Freq ~ ., data = tbldat)
      names(dimnames(crosstab))[z - 1] <- ""
    }

    #Add margins if required, adding only those margins appropriate to user request
    if (addmargins==TRUE) {

      vars <- c(row.vars,col.vars)

      if (length(type)==1) {
      if (type=="row.pct") 
        { crosstab <- addmargins(crosstab,margin=c(vars[n.vars]))
          tbl <- addmargins(tbl,margin=c(vars[n.vars]))
        }
      else 
        { if (type=="column.pct") 
            { crosstab <- addmargins(crosstab,margin=c(vars[n.row.vars]))
              tbl <- addmargins(tbl,margin=c(vars[n.row.vars]))
            }
          else 
            { if (type=="joint.pct") 
              { crosstab <- addmargins(crosstab,margin=c(vars[(n.row.vars)],vars[n.vars])) 
                tbl <- addmargins(tbl,margin=c(vars[(n.row.vars)],vars[n.vars])) 
              }
              else #must be total.pct OR frequency
                { crosstab <- addmargins(crosstab)
                  tbl <- addmargins(tbl)
                }
            }
  	     } 
      }

      #If more than one table type requested, only adding row totals makes any sense...
      if (length(type)>1) {
        crosstab <- addmargins(crosstab,margin=c(vars[n.vars]))
        tbl <- addmargins(tbl,margin=c(vars[n.vars]))
      }

    }

    #If subtotals not required, and total vars > 2, create dataframe version of table, with relevent
    #subtotal rows / cols dropped [Subtotals only present in tables with > 2 cross-classified vars]
    t1 <- NULL
    if ( (subtotals==FALSE) & (n.vars>2) )  {

    #Create version of crosstab in ftable format
    t1 <- crosstab 
    t1 <- ftable(t1,row.vars=row.vars,col.vars=col.vars)

    #Convert to a dataframe
    t1 <- as.data.frame(format(t1),stringsAsFactors=FALSE)

>>>>

---
title: Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Subjective Wellbeing
author:
  - name: Beatriz Mella Lira
    email: beatriz.lira.14@ucl.ac.uk
    affiliation: University College London
    footnote: Corresponding Author
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
address:
  - code: University College London
    address: Bartlett School of Planning, Sixth Floor, Central House, 14 Upper Woburn Place, London, WC1H0NN UK
  - code: McMaster University
    address: School of Geography and Earth Sciences, 1280 Main St W, Hamilton, Ontario, L8S 1S4 Canada
abstract: |
  Subjective wellbeing is a topic that has attracted considerable attention in recent years due to the way it correlates with health. From a transportation perspective, there is a burgeoning literature on the way travel can impact subjective wellbeing, and how this, in turn, can influence behavior. After decades of fascination with the automobile, subjective wellbeing is increasingly recognized as an alternative approach to assess transportation, and to gain valuable insights to attract and retain active travelers and users of public transportation. The objective of this paper is to analyze a number of affective values associated with subjective wellbeing and the modes of transportation that people commonly use. In particular, we are interested in the potential for dissonance with respect to the primary mode of travel. The study is based on data collected from a sample of travellers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name the mode or modes that they associate with the affective values of freedom, enjoyment, happiness, poverty, luxury and status. The results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, public transportation is most commonly associated with poverty, and the automobil is most commonly associated with luxury and status.

@@
